---
title: "Final Project Data Analyst Career"
output: html_document
author: "Alexander Nicolás Tisko"
url: {https://github.com/B1ggieSmalls/Data-Analysis-Final-Project-}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis Final Project - Coursera 

This is my final Google career project. The **objective** is to identify trends so marketing company can work efficient in a campaign for new clients. We are going to use a free and public data offered by [Möbius](https://www.kaggle.com/arashnic). The data set is: https://www.kaggle.com/arashnic/fitbit. 

Some of the data was modified by hand on excel spreadsheet and final details were done in R. 

## Packages installed and libraries

```{r message=FALSE}
  library(tidyverse)
  library(here)
  library(ggplot2)
  library(skimr)
  library(janitor)
  library(readxl)
```

### Importing first data set:

```{r message=FALSE}
dailiy_Activity <- read_excel("dailiy_Activity.xlsx")
```

#### Cleaning and analyzing data

```{r message=FALSE}
  ##Cleaning column names
  da <- clean_names(dailiy_Activity)
```

```{r}
da %>%
    select(total_steps,
           total_distance,
           very_active_distance,
           moderately_active_distance,
           light_active_distance,
           very_active_minutes,
           lightly_active_minutes,
           fairly_active_minutes,
           calories) %>%
    summary()
```

### Light exercise is the most done

We can see trends in light distance and minutes means, the values are bigger compared with the other columns This means that people use to do more exercise in a lightly way. Pay attention in the **means data** of the next columns and compare it with the columns at the top. 

```{r echo=FALSE}
da %>%
  select(light_active_distance,
         lightly_active_minutes) %>%
  summary()
```

### You burn more calories doing an active excercise

In the next graph, we can see an **important trend**. People who do more active exercise, burn more calories than lighter and fairer exercises. 

```{r echo=FALSE}
ggplot(data = da, aes(x = calories, y = total_activity_minutes)) +
      geom_point(aes(alpha = lightly_active_minutes), color = "red") +
      geom_point(aes(alpha = very_active_minutes), color = "blue") +
      geom_point(aes(alpha = fairly_active_minutes), color = "green") +
      theme(axis.title.x = element_text()) +
      labs(x = "Calories", y = "Total activity (mins)") +
      annotate("text", x = 680, y = 530, label = "Red = Lightly active mins", color = "red") +
      annotate("text", x = 645, y = 500, label = "Blue = Very active mins", color = "blue") +
      annotate("text", x = 705, y = 470, label = "Green = Fairly active mins", color = "green") +
      theme(legend.position='none') 
```

### How many time do I need to burn calories? And which excercise is the bestone?
We can see, that if you do light activity u need more time to burn calories. But if you do an extensive activity, u need less time. So, if you are exercised this is an excellent notice!

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
   ggplot(data = da) +
      geom_point(mapping = aes(x = calories, y = total_activity_minutes)) +
      geom_smooth(mapping = aes (x = calories, y = very_active_minutes), color = "blue") +
      geom_smooth(mapping = aes (x = calories, y = lightly_active_minutes), color = "red") +
      labs(x = "Calories", y = "Total activity (mins)") +
      theme(axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold")) +
      annotate("text", x = 500, y = 530, label = "Blue = Very active mins", color = "blue") +
      annotate("text", x = 550, y = 500, label = "Red = Lightly active mins", color = "red") 
```

### Humans spend more time as active or sedentary?

We can see that humans spend more time sedentary than active. When I saw this data, I thought in the time I spend working, studying, playing pc (sedentary) and the time I am doing exercise or moving. 

**Total minutes = sedentary + active**

```{r echo=FALSE, message=FALSE}
ggplot(data = da) +
      geom_point(mapping = aes(x = calories, y = total_minutes)) +
      geom_smooth(mapping = aes (x = calories, y = sedentary_minutes), color = "red") + 
      geom_smooth(mapping = aes (x = calories, y = total_activity_minutes), color = "yellow") +
      labs(x = "Calories", y = "Total minutes") +
      annotate("text", x = 3500, y = 100, label = "Red = Sedentary active mins", color = "blue") +
      annotate("text", x = 3435, y = 40, label = "Yellow = Total active mins", color = "blue") 

```

## Importing new data set

```{r message=FALSE}
sleep_Day <- read_excel("sleep_Day.xlsx")
```

## Cleaning and analyzing it

```{r}
 sl <- clean_names(sleep_Day) %>%
      select(-id_len)
  ##Only for future datasets merges & analysis, I will change the column named as SleepDay for activity_date
  sld <- rename(sl, activity_date = sleep_day)
```

```{r}
    sld %>%
      select(hour_and_minutes_asleep,
             hours_and_minutos_in_bed,
             total_minutes_asleep,
             total_time_in_bed) %>%
      summary()
```
## Minutes asleep vs in bed

We found a trend that humans spend more time in bed than only sleeping. When I saw this data, I remembered the time I spend maybe with the phone in the bed or the computer. 

```{r echo=FALSE}
ggplot(data = sld) +
      geom_point(mapping = aes(x = total_time_in_bed, y = total_minutes_asleep), color = "purple") +
      labs(x = "Total time in bed", y = "Total minutes asleep")
      
```

## Merge time!

We arrive to the merge time. We are going to merge the data sets, so we can find new trends and compare both data. How are we going to merge? We are going to use the inner join way, because we need only the information that match in both datasets. Ok, let's do it!

```{r}
da_sl <- merge(da, sld, by = c("id","activity_date"))
head(da_sl)
```

## Cleaning and analyzing 

***Important:*** I want to know the people who sleep more and less than 8 hours and compare the data so maybe we can find new trends. 

```{r}
 filter_8hours <- da_sl %>%
        filter(total_minutes_asleep >= 480)
    ##Conclusion: I had 413 rows in the data set and 117 people slept more than 8 hours. 
      filter_less8hourse <- da_sl %>%
        filter(total_minutes_asleep < 480)
    ##Conclusion_ From 413 rows data, 296 slept less than 8 hours.

```

```{r echo=FALSE}
    filter_hours_pie <- c(117, 296)
    piepercent <- round(100*filter_hours_pie/sum(filter_hours_pie), 1)
    pie(filter_hours_pie, labels = piepercent, col = rainbow(length(filter_hours_pie)))
    legend("topright", c("8 hourse or more", "Less than 8 hours"), cex = 0.8,
           fill = rainbow(length(filter_hours_pie)))
```

### Important trends from last merge and filter

We are going to select only the columns we need for analysis and show a summary of both data sets ( x < 8 && x >= 8). In these summaries, we can compare the information and see some trends. ***We are going to focus in means values.*** 

I created 2 data sets. One for the people who sleep equal or more than 8 hours, and other one for people who sleep less than 8 hours. We can see from my original data set, 117 people sleep 8 or more hours and 296 less than 8 hours.

So, if we only focus on means values, we are going to face an issue, cause is not going to be the same for both data sets. We need to be partial. In this case, I am going to pick a 117 rows sample from my 296 rows data set. Now, both data sets are going to have 117 rows and we can compare means.  


```{r echo=FALSE}
      ##fl8 = filter_less8hours
      select_fl8 <- filter_less8hourse %>%
        select(total_steps,
               total_distance,
               very_active_distance,
               moderately_active_distance,
               light_active_distance,
               very_active_minutes,
               fairly_active_minutes,
               lightly_active_minutes,
               total_activity_minutes,
               sedentary_minutes, 
               calories,
               total_minutes_asleep,
               total_time_in_bed)
      
      ##8h = 8hours
      select_8h <- filter_8hours %>%
        select(
          total_steps,
          total_distance,
          very_active_distance,
          moderately_active_distance,
          light_active_distance,
          very_active_minutes,
          fairly_active_minutes,
          lightly_active_minutes,
          total_activity_minutes,
          sedentary_minutes, 
          calories,
          total_minutes_asleep,
          total_time_in_bed)
      
     
```



```{r}
##I already created both data sets, this is only the random formula
random_selectfl8 <- select_fl8 %>%
        sample_n(117, replace = FALSE)
```

### Less than 8 hours

```{r echo=FALSE}
summary(random_selectfl8)
```

### 8 Hours or more

```{r echo=FALSE}
summary(select_8h)
```

### Conclusion 

We can see a trend that people who sleep -8 hours use to do more active, lightly and moderately exercise and also used to spend more time doing this. 

## To complete...

Finally, after analyze data with numbers, it is time to see visual trends. We are going to merge the two 117 rows data sets, the people who sleep equal or more than 8 hours and less than 8 hours, so we have a new data set with 234 rows and with the information combine.  

In this case, we are going to merge with outter join, cause we need both data and not only the ones who are the same. 

```{r}
merge_8h_fl8 <- merge(select_8h, random_selectfl8, all = TRUE)
      
eight_hours <- merge_8h_fl8 %>%
  mutate(eight_hours_or_not = if_else(total_minutes_asleep >= 480, 'TRUE', 'FALSE'))
glimpse(eight_hours)
```

## Interesting graphs

### Sedentary Minutes vs Tota Activity Minutes

We can find some imbalance in sedentary minutes comparing to activity minutes in people who sleep less than 8 hours. On people who sleep 8 hours or more we find a balance and the sedentary minutes are not so high, they trend to be balance and center.  


```{r echo=FALSE}
 label_filter_change = as_labeller(c(`FALSE` = "Less than 8 hours", `TRUE` = "8 hours or more"))
  ggplot(data = eight_hours) +
    geom_point(mapping = aes(x = total_activity_minutes, y = sedentary_minutes, color = calories)) +
    facet_wrap(~eight_hours_or_not, labeller = label_filter_change) +
    labs(x = "Total Activity Minutes", y = "Sedentary Minutes")
```

### Balance between the tota distance and the calories. 

We can see a balance also in people who sleep 8 hours or more according to the different type of activities. 

***Blue*** = Very Active
***Green*** = Moderately active
***Red*** = Light Activity

```{r echo=FALSE, message=FALSE}
    ggplot(data = eight_hours, aes(x = total_distance, y = calories)) +
      geom_smooth(mapping = aes(x = very_active_distance, y = calories), se = FALSE) +
      geom_smooth(mapping = aes(x = light_active_distance, y = calories), color = "red", se = FALSE) +
      geom_smooth(mapping = aes(x = moderately_active_distance, y = calories), color = "green", se = FALSE) +
      facet_wrap(~eight_hours_or_not, labeller = label_filter_change) +
      labs(x ="Total Distance", y = "Calories") 
```

## Conclusion

We can see in the summaries that people who sleep less has higher mean values, almost in all columns, but according to the graphs, we can see that the people who sleep 8 or more hours are more balance than the people who sleep less. This means, we can find a breakeven in +8 hours, and in - 8 hours we find lot of peaks values.  

## Importing more data 

We are going to import the weight (kg) values, including bmi also (body mass index).   

```{r}
weight_Log_Info <- read_excel("weight_Log_Info.xlsx")
```

## Cleaning and analyzing 

```{r}
head(weight_Log_Info)

weightt <- weight_Log_Info %>%
      clean_names() %>%
      rename(activity_date = x3) 

weight <- weightt %>%
      select(id,
             activity_date,
             weight_kg,
             fat,
             bmi,)
colnames(weight)
```

## Merging da_sl data set with the new weight data

It is going to be merge with inner join, we need only the information that match. 

```{r}
  da_sl_we <- merge(da_sl, weight, by = c("id", "activity_date"))
```

We can see a trend when weight increment, the bmi also increment. 

```{r echo=FALSE}

ggplot(data = da_sl_we) +
  geom_line(mapping = aes(x = weight_kg, y = bmi, ), color = "blue") +
  labs(x = "Weight (kg)", y = "BMI")
```

## Importing new data sets, cleaning, analyzing and merging

Now, we want to analyze the data sets per hour and merge them for trends findings. We are going to import data sets, clean it, analyze it so we can see visual trends. 

```{r}
    hourly_Intensities <- read_excel("hourly_Intensities.xlsx")
    hourly_Calories <- read_excel("hourly_Calories.xlsx")
    hourly_Steps <- read_excel("hourly_Steps.xlsx") 
```

**hi =hourly_Itennsities**
**hc = hourly_Calories**
**hs = hourly_Steps**

#### Cleaning

```{r}
hi <- hourly_Intensities %>%
  clean_names()
    
hc <- hourly_Calories %>%
  clean_names()
      
hs <- hourly_Steps %>%
  clean_names()
```

#### Merging

```{r}
hi_hc <- merge(hi, hc, by = c("id", "activity_date"))
```

### Cleaning and analyzing

This merge is so extensive, like 254979 observations, so we are going to take a sample, according to the poblation so we can analyze it and do future merges. 

```{r}
random_hi_hc <- hi_hc %>%
            sample_n(1063, replace = FALSE)
```

### Merging with hours_steps

```{r}
hi_hc_hs <- merge(random_hi_hc, hs, by = c("id", "activity_date"))
```

This new merge has duplicates on same dates and minutes, so we are going to clean this duplicates. 

```{r}
  hi_hc_hs_nodup <- hi_hc_hs %>%
        distinct(id, activity_date, .keep_all = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
